---
title: "STA304 A2"
author: "Zifan Pan"
date: '1004067332'
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## **Assignment 2**\
### **Question 3**\
\
```{r input info, echo=TRUE}
# Name Brand I stratum 1:
N1 = 120
lastyrSale1 = 24500
popMu1 = lastyrSale1/N1
# the sample values are:
x1 <- c(204, 143, 82, 256, 275, 198)
y1 <- c(210, 160, 75, 280, 300, 190)
# number of samples:
n1 = length(y1)


# Name Brand II stratum 2:
N2 = 180
lastyrSale2 = 21200
popMu2 = lastyrSale2/N2
# the sample values are:
x2 <- c(137, 189, 119, 63, 103, 107, 159, 63, 87)
y2 <- c(150, 200, 125, 60, 110, 100, 180, 75, 90)
# number of samples:
n2 = length(y2)


# total population size
N = N1+N2
# weight
W1 = N1/N
W2 = N2/N
# fractions
f1 = n1/N1
f2 = n2/N2
# sample mean for brand I
y1_bar = mean(y1)
x1_bar = mean(x1)
# sample mean for Brand II
y2_bar = mean(y2)
x2_bar = mean(x2)
# sample variance for brand I
y1_s2 = var(y1)
x1_s2 = var(x1)
# sample variance for Brand II
y2_s2 = var(y2)
x2_s2 = var(x2)
# coefficients between x and y (brand I)
corr1 = cor(y1, x1)
# coefficients between x and y (brand II)
corr2 = cor(y2, x2)

```
\
#### **Part 1**\
```{r part1_1}
tau_hat = N1*y1_bar+N2*y2_bar
tau_hat
```
\
From the output above, the basic estimate of total potential sales is:\
<center>$\hat\tau=\hat\tau_1+\hat\tau_2=(N_1\bar y_1)+(N_2\bar y_2)=46100$</center>\
\
```{r part1_2}
var_hat_mu_hat_1 = (1-f1)*(y1_s2/n1)
var_hat_mu_hat_2 = (1-f2)*(y2_s2/n2)
var_hat_tau_hat = N1^2*var_hat_mu_hat_1+N2^2*var_hat_mu_hat_2
var_hat_tau_hat
```
\
From the output above, the estimated variance is:\
<center>$\hat V(\hat\tau)=\hat V(\hat\tau_1)+\hat V(\hat\tau_2)= N_1^2\times(1-n_1/N_1)\times s^2_{y_1}/n_1 +N_2^2\times(1-n_2/N_2)\times s^2_{y_2}/n_2=23075975$ </center>\
\
#### **Part 2**\
**(a)**\
```{r part2_1}
# ratio
r1 = y1_bar/x1_bar
r2 = y2_bar/x2_bar
mu_y_hat_sr_1 = W1*r1*popMu1
mu_y_hat_sr_2 = W2*r2*popMu2
mu_y_hat_sr = mu_y_hat_sr_1+mu_y_hat_sr_2
tau_y_hat_sr = N*mu_y_hat_sr
tau_y_hat_sr
```
From the output above, the ratio estimate of the total potential sales is:\
<center>$\hat\tau_{r,sr}= N\times\hat\mu_{y,SR}=N[W_1\times(\bar y_1/\bar x_1)\times \mu_{x_1} + W_2\times (\bar y_2/\bar x_2)\times \mu_{x_2}]=48206.45$</center>\
\
```{r part2_2}
r1_s2 = var(y1-r1*x1)
r2_s2 = var(y2-r2*x2)
var_hat_mu_r_hat_1 = (1-f1)*(r1_s2/n1)
var_hat_mu_r_hat_2 = (1-f2)*(r2_s2/n2)
weight_var_1 = W1^2*var_hat_mu_r_hat_1
weight_var_2 = W2^2*var_hat_mu_r_hat_2
weight_var = weight_var_1+weight_var_2
var_tau_hat_r_sr = weight_var*N^2
var_tau_hat_r_sr
```
From the output above, the estimated variance is:\
<center>$\hat V(\hat\tau_{r,sr})=N^2[W_1^2\times(1-n_1/N_1)\times(s^{r_1}/n_1)+W_2^2\times(1-n_2/N_2)\times(s^{r_2}/n_2)]= 564612.8 $</center>\
\
**(b)**:\
```{r part2_b_1}
# cross products between x, y:
sx1y1 = sum((x1-mean(x1))*(y1-mean(y1)))/(n1-1)
sx2y2 = sum((x2-mean(x2))*(y2-mean(y2)))/(n2-1)
# regression coefficient for brand I:
b1 = sx1y1/x1_s2
b2 = sx2y2/x2_s2
a1 = y1_bar-b1*x1_bar
a2 = y2_bar-b2*x2_bar
# output
b1
b2
a1
a2
```
From the R output above, the regression coefficient for strata 1 is:\
<center>$b_1=s_{x_1y_1}/s^2_{x_1}=1.1370$</center>\
<center>$a_1=\bar y_1-b_1\bar x_1=-16.93784$</center>\
From the R output above, the regression coefficient for strata 2 is:\
<center>$b_2=s_{x_2y_2}/s^2_{x_2}=1.1065$</center>\
<center>$a_2=\bar y_2-b_2\bar x_2=-5.157056$</center>\
\
```{r part2_b_2}
mu_l_hat_1 = y1_bar+b1*(popMu1-x1_bar)
mu_l_hat_2 = y2_bar+b2*(popMu2-x2_bar)
mu_l_hat_sr_1 = W1*mu_l_hat_1
mu_l_hat_sr_2 = W2*mu_l_hat_2
mu_l_hat_sr = mu_l_hat_sr_1 + mu_l_hat_sr_2
tau_l_hat_sr = N*mu_l_hat_sr
tau_l_hat_sr
```
From the R output above, the regression estimator is:\
<center>$\hat\tau_{L,sr}=N\times\hat\mu_{L,sr}=N[W_1\times(\bar y_1+b_1(\mu_{x1}-\bar x_1))+ W_2\times(\bar y_2+b_2(\mu_{x2}-\bar x_2))]=48353.87$</center>\
\
```{r part2_b_3}
# y hat:
y1_hat = a1+b1*x1
y2_hat = a2+b2*x2

e1 = y1-y1_hat
e2 = y2-y2_hat

MSE_1 = sum(e1^2)/(n1-2)
MSE_2 = sum(e2^2)/(n2-2)
MSE_1
MSE_2

```
Before calculating the estimated variance, we first need to find the MSE of each stratum, based on the R output above, we can get:\
<center>$MSE_1=[1/(n_1-2)*\sum^{n_1}_{i=1}(y_{1,i}-(a_1+b_1x_{1,i}))^2]=154.0006$</center>\
<center>$MSE_2=[1/(n_2-2)*\sum^{n_2}_{i=1}(y_{2,i}-(a_2+b_2x_{2,i}))^2]=60.51966$</center>\
\
```{r part2_b_4}
varHat_muLhat_sr1 = W1^2*(1-f1)*(MSE_1/n1)
varHat_muLhat_sr2 = W2^2*(1-f2)*(MSE_2/n2)
varHat_muLhat_sr = varHat_muLhat_sr1 + varHat_muLhat_sr2
varHat_tauLhat_sr = N^2*varHat_muLhat_sr
varHat_tauLhat_sr
```
From the output above, the estimated variance is:\
<center>$\hat V(\hat\tau_{L,sr})=N^2[W_1^2\times(1-n_1/N_1)\times MSE_1/n_1 + W_2^2\times(1-n_2/N_2)\times MSE_2/n_2]=558098.7$</center>\
\
**(c)**:\
```{r part2_c}
effic_rb = var_tau_hat_r_sr/var_hat_tau_hat
effic_rb
```
The R output calculated above is the relative efficiency of ratio estimator ($\hat\tau_{r,sr}$) to the basic one ($\hat\tau$), which is 0.024 less than 1. Which means $\hat V(\hat\tau_{r,sr})<\hat V(\hat\tau)$, so that ratio estimator ($\hat\tau_{r,sr}$) is preferred.\
```{r part2_c_2}
effic_rl = var_tau_hat_r_sr/varHat_tauLhat_sr
effic_rl
```
The R output calculated above is the relative efficiency of ratio estimator ($\hat\tau_{r,sr}$) to the regression one ($\hat\tau_{L,sr}$), which is 1.011 approximately equals to 1. Which means $\hat V(\hat\tau_{r,sr})\approx\hat V(\hat\tau_{L,sr})$, so that both estimators are equally preferred.\
```{r part_c_2}
effic_bl = var_hat_tau_hat/varHat_tauLhat_sr
effic_bl
```
The R output calculated above is the relative efficiency of basic estimator ($\hat\tau$) to the regression one ($\hat\tau_{L,sr}$), which is 41.347 greater than 1. Which means $\hat V(\hat\tau_{L,sr})<\hat V(\hat\tau)$, so that regression estimator ($\hat\tau_{L,sr}$) is preferred.\
\
**In Conclusion,** we would recommend both separate ratio and regression estimators.\
\
\
#### **Part 3**\
**(a)**\
```{r part3_a_1}
popMu = W1*popMu1+W2*popMu2
popMu
y_bar_st = W1*y1_bar + W2*y2_bar
y_bar_st
x_bar_st = W1*x1_bar + W2*x2_bar
x_bar_st

# combined ratio estimate of mean:
mu_hat_cr = (y_bar_st/x_bar_st)*popMu
mu_hat_cr
```
Based on the R output, we obtain that:\
<center>$\mu_x= W_1*\mu_{x_1}+W_2*\mu_{x_2}=152.3333$</center>\
<center>$\bar y_{st}= W_1*\bar y_{1}+W_2*\bar y_2=153.6667$</center>\
<center>$\bar x_{st}= W_1*\bar x_{1}+W_2*\bar x_2=145.6667$</center>\
The combined ratio estimate of $\mu$ is:\
<center>$\hat\mu_{cr}=(\bar y_{st}/\bar x_{st})\times\mu_x=160.6995$</center>\
\
```{r part3_a_2}
# ratio (CR):
r_cr = (y_bar_st/x_bar_st)
# for stratum 1:
s2_cr1 = sum((y1-r_cr*x1)^2)/(n1-1)
s2_cr2 = sum((y2-r_cr*x2)^2)/(n2-1)
s2_cr1
s2_cr2
# estimated variance:
var_hat_mu_hat_cr = W1^2*(1-f1)*(s2_cr1/n1) + W2^2*(1-f2)*(s2_cr2/n2)
var_hat_mu_hat_cr
```
From the output above, the estimated variance is:\
<center>$\hat V(\hat\mu_{cr})=W_1^2\times(1-n_1/N_1)\times(s^2_{cr,1}/n_1) + W_2^2\times(1-n_2/N_2)\times(s^2_{cr,2}/n_2)=6.25$</center>\
<center>where $s^2_{cr,1}=[\sum^{n_1}_{j=1}(y_j-\hat r_{cr}x_j)^2/(n_1-1)]=159.2125$,</center>\
<center>and $s^2_{cr,2}=[\sum^{n_2}_{j=1}(y_j-\hat r_{cr}x_j)^2/(n_2-1)]=58.3233$</center>\
\
**(b)**\
```{r part3_b_1}
# cross products between x, y:
sx1y1 = sum((x1-mean(x1))*(y1-mean(y1)))/(n1-1)
sx2y2 = sum((x2-mean(x2))*(y2-mean(y2)))/(n2-1)
# regression coefficient
b1 = sx1y1/x1_s2
b2 = sx2y2/x2_s2
a1 = y1_bar-b1*x1_bar
a2 = y2_bar-b2*x2_bar
# output
b1
b2
a1
a2
```
From the R output above, the regression coefficient for strata 1 is:\
<center>$b_1=s_{x_1y_1}/s^2_{x_1}=1.1370$</center>\
<center>$a_1=\bar y_1-b_1\bar x_1=-16.93784$</center>\
From the R output above, the regression coefficient for strata 2 is:\
<center>$b_2=s_{x_2y_2}/s^2_{x_2}=1.1065$</center>\
<center>$a_2=\bar y_2-b_2\bar x_2=-5.157056$</center>\
\
```{r part3_b_2}
mu_l_hat_1 = y1_bar+b1*(popMu1-x1_bar)
mu_l_hat_2 = y2_bar+b2*(popMu2-x2_bar)
mu_l_hat_sr_1 = W1*mu_l_hat_1
mu_l_hat_sr_2 = W2*mu_l_hat_2
mu_l_hat_sr = mu_l_hat_sr_1 + mu_l_hat_sr_2
tau_l_hat_sr = N*mu_l_hat_sr
tau_l_hat_sr
```
From the R output above, the regression estimator is:\
<center>$\hat\tau_{L,sr}=N\times\hat\mu_{L,sr}=N[W_1\times(\bar y_1+b_1(\mu_{x1}-\bar x_1))+ W_2\times(\bar y_2+b_2(\mu_{x2}-\bar x_2))]=48353.87$</center>\
\
```{r part3_b_3}
# y hat:
y1_hat = a1+b1*x1
y2_hat = a2+b2*x2

e1 = y1-y1_hat
e2 = y2-y2_hat

MSE_1 = sum(e1^2)/(n1-2)
MSE_2 = sum(e2^2)/(n2-2)
MSE_1
MSE_2
```
Before calculating the estimated variance, we first need to find the MSE of each stratum, based on the R output above, we can get:\
<center>$MSE_1=[1/(n_1-2)*\sum^{n_1}_{i=1}(y_{1,i}-(a_1+b_1x_{1,i}))^2]=154.0006$</center>\
<center>$MSE_2=[1/(n_2-2)*\sum^{n_2}_{i=1}(y_{2,i}-(a_2+b_2x_{2,i}))^2]=60.51966$</center>\
\
```{r part3_b_4}
varHat_muLhat_sr1 = W1^2*(1-f1)*(MSE_1/n1)
varHat_muLhat_sr2 = W2^2*(1-f2)*(MSE_2/n2)
varHat_muLhat_sr = varHat_muLhat_sr1 + varHat_muLhat_sr2
varHat_tauLhat_sr = N^2*varHat_muLhat_sr
varHat_tauLhat_sr
```
From the output above, the estimated variance is:\
<center>$\hat V(\hat\tau_{L,sr})=N^2[W_1^2\times(1-n_1/N_1)\times MSE_1/n_1 + W_2^2\times(1-n_2/N_2)\times MSE_2/n_2]=558098.7$</center>\
\
**(c)**\
```{r part3_c_1}
var_hat_tau_hat_cr = N^2*var_hat_mu_hat_cr
effic_rb2 = var_hat_tau_hat_cr/var_hat_tau_hat
effic_rb2
```
The R output calculated above is the relative efficiency of ratio estimator ($\hat\tau_{r,sr}$) to the basic one ($\hat\tau$), which is 0.024 less than 1. Which means $\hat V(\hat\tau_{r,sr})<\hat V(\hat\tau)$, so that ratio estimator ($\hat\tau_{r,sr}$) is preferred.\
```{r part3_c_2}
effic_rl2 = var_hat_tau_hat_cr/varHat_tauLhat_sr
effic_rl2
```
The R output calculated above is the relative efficiency of ratio estimator ($\hat\tau_{r,sr}$) to the regression one ($\hat\tau_{L,sr}$), which is 1.0078 approximately equals to 1. Which means $\hat V(\hat\tau_{r,sr})\approx\hat V(\hat\tau_{L,sr})$, so that both estimators are equally preferred.\
```{r part3_c_3}
effic_bl2 = var_hat_tau_hat/varHat_tauLhat_sr
effic_bl2
```
The R output calculated above is the relative efficiency of basic estimator ($\hat\tau$) to the regression one ($\hat\tau_{L,sr}$), which is 41.347 greater than 1. Which means $\hat V(\hat\tau_{L,sr})<\hat V(\hat\tau)$, so that regression estimator ($\hat\tau_{L,sr}$) is preferred.\
\
**In Conclusion,** we would recommend both combined ratio and regression estimators.\
\
\
**END of Question 3.**













## **Q4:**
```{r}
# load data
mydata<-read.csv("StudentsMarks.csv")
# use x to record term test 1 marks,
# after removing students who missed the test.
x<-na.omit(mydata$Test.1)
N<-length(x)
N
```

